---
title : 'W271 Assignment 3'
output: 
  pdf_document: 
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r load packages, message=FALSE}
library(tidyverse)
library(car)
library(sandwich)
library(lmtest)
library(knitr)
library(Hmisc)
library(gridExtra)
library(stargazer)
library(mcprofile)
```


# Admission Data: Binary Logistic Regression 
(One point per question/sub-question. Eight points total.)

The data set *"admissions.csv"* contains a small sample of graduate school admission data from a university. The variables are specified below:

  1. `admit` - the dependent variable that takes two values: $0,1$ where $1$ denotes *admitted* and $0$ denotes *not admitted*
  2. `gre`  - GRE score
  3. `gpa`  - College GPA
  4. `rank` - rank in college major

Suppose you are hired by the University's Admission Committee and are charged to analyze this data to quantify the effect of GRE, GPA, and college rank on admission probability. We will conduct this analysis by answering the following questions:

```{r read admission data, message=FALSE}
admission <- read_csv('./data/admissions.csv')
```

## Examine the data and conduct an EDA
Examine the data and conduct EDA. Are there any points that are strange, or outlying? Are there any features of the data that affect how you will analyze it?

```{r conduct EDA}

```

> 'Fill this in: What do you note about this data?' 

## Estimate a logistic regression
Estimate the following binary logistic regressions: 

$$
  \begin{aligned}
    Y = \beta_{0} & + \beta_{1} GRE + \beta_{2} GPA + \beta_{3} RANK + e & \text{(Model 1)} \\ 
    Y = \beta_{0} & + \beta_{1} GRE + \beta_{2} GPA + \beta_{3} RANK & \text{(Model 2)} \\ 
      & + \beta_{4} GRE^2 + \beta_{5} GPA^2 + e\\
    Y = \beta_{0} &+ \beta_{1} GRE + \beta_{2} GPA + \beta_{3} RANK & \text{(Model 3)} \\ 
      & + \beta_{4} GRE^2 + \beta_{5} GPA^2 \\ 
      & + {\beta_6} GRE \times GPA + e
  \end{aligned}
$$

where $GRE \times GPA$ denotes the interaction between `gre` and `gpa` variables.

```{r estimate logistic regression with interaction}
model_admission_1 <- 'fill this in'
model_admission_2 <- 'fill this in'
model_admission_3 <- 'fill this in'

## display estimated model in a table

#stargazer(model_admission_1, model_admission_2, model_admission_3, type = "text", omit.stat = "f",
#                    star.cutoffs = c(0.05, 0.01, 0.001), title = "Table 1: The estimated relationship between Admission and GRA, #GPA, and Students' Rank")

```

## Test hypotheses

### Linear effect: class rank
Using `model_admission_1`, test the hypothesis that class rank has no effect on admission using a likelihood ratio test. Suppose that someone asks, "Are we willing to assume that there is a *linear* effect of class rank as we have in `model_admission_1`?"   

```{r rank effect}

```

> 'Fill this in: What do you observe?'

### Linear effect: GRE
Test the hypothesis that $\beta_{1} = 0$ in `model_admission_2` using a likelihood ratio test. Interpret what this test result means in the context of a model like what you have estimated in `model_admission_2`. 

Then, test the same hypothesis in `model_admission_3` using a likelihood ratio test. Interpret what this test result means in the context of a model like what you have estimated in `model_admission_3`. 

```{r GRE LRT}

```

### Total effect: GRE
Test the hypothesis that $GRE$ has no effect on the likelihood of admission, in a model of admissions defined in `model_admission_3`, using a likelihood ratio test. 

```{r test for no GRE effect}

```

## Interpret an effect

Using the entire model, make predictions about how the likelihood of being admitted changes for someone with a $GPA = 3.0$ compared to someone with a $GPA = 4.0$ both with $GRE=600$. 

```{r GPA impact on admission}

```


## Construct a confidence interval 
Construct the 95% Profile LR confidence interval for the admission probability for the students with the following profile using **model_admission_3**: 

- $GPA  = 3.3$; 
- $GRE  = 720$; and, 
- $rank = 1$


- $GPA  = 2.5$; 
- $GRE  = 790$; and,  
- $rank = 4$. 

```{r construct confidence interval for admission}
admission_probabilty_ci <- 'fill this in'
```

> 'Fill this in: Use inline code evaluation to put all of your written content into the printed page.'

Are the prediction intervals for these two predictions the same? Why or why not? What about the data leads to this similarity or difference? 

> 'Fill this in'
