# (10 points) ARIMA model

Consider `fma::sheep`, the sheep population of England and Wales from 1867--1939. :sheep: 

```{r load sheep data, message=FALSE, warning=FALSE}
#install.packages('fma')
library(fma)
head(fma::sheep)
```

## Time series plot
Produce a time plot of the time series, and comment on what you observe.

```{r produce a time series plot of sheep}
sheep_plot <- 'fill this in'
```

> 'Fill this in' 

## Fit a model 

Assume you decide to fit the following model: 

$$
  y_{t} = y_{t-1} + \phi_1(y_{t-1}-y_{t-2}) + \phi_2(y_{t-2}-y_{t-3}) + \phi_3(y_{t-3}-y_{t-4}) + \epsilon_t,
$$ 

where $\epsilon_t$ is a white noise series.

### Model type
What sort of ARIMA model is this (i.e., what are $p$, $d$, and $q$)?

> 'Fill this in.'

### Back to the future 

Express this ARIMA model using backshift operator notation.

> 'Fill this in.'

## Is this model appropriate? 

Examine the ACF and PACF of the differenced data. Evaluate whether this model is appropriate. 

```{r}

```

> 'Fill this in.'

## Forecasts, by hand! 
The last five values of the series are given below:

| Year              | 1935 | 1936 | 1937 | 1938 | 1939 |
|:------------------|-----:|-----:|-----:|-----:|-----:|
| Millions of sheep | 1648 | 1665 | 1627 | 1791 | 1797 |

The estimated parameters are: 

- $\phi_1 =  0.42$; 
- $\phi_2 = -0.20$;  and, 
- $\phi_3 = -0.30$.

Without using the forecast function, calculate forecasts for the next three years (1940--1942).

```{r create forcasts by hand}
## you can use R as a calculator here

```

## Interpret roots
Find the roots of your model's characteristic equation. Is this process stationary?.

```{r}

```

> 'Fill this in.'