# (10 points) Time Series Linear Model and Cointegration

Daily electricity demand and temperature (in degrees Celsius) is recorded in `./data/temperature_demand.csv`. Please work through the following questions to build a time series linear model against this data.

```{r load packages for time series linear model, warning=FALSE, message=FALSE}
library(tidyverse)
```

```{r load temperature data, message=FALSE} 
temperature <- read_csv('./data/temperature_demand.csv') %>% 
  rename(
     'index'       = '...1',
     'demand'      = 'Demand', 
     'work_day'    = 'WorkDay',
     'temperature' = 'Temperature'
  )
#glimpse(temperature)
```

## Plot electricity

Plot electricity demand and temperature as time series. Is there any correlation between these to variables? If yes, Do you think is it a spurious correlation?

```{r plot electricity and temperature}
#'fill this in'

```

> 'Fill this in

## Cointegration test

Use the Engle-Granger test to check for cointegration. What do you conclude? 

```{r Cointegration test}
#'fill this in'

``` 

## Fit Model

Based on cointegration test,fit a regression model for demand with temperature as an explanatory variable(or their first difference).

```{r Fit model}
#'fill this in'

``` 


## Residuals Plot

Produce a residual plot of the estimated model in pervious part. Is the model adequate? Describe any outliers or influential observations, and discuss how the model could be improved.

```{r plot residuals}
#Fill this in

```

> 'Fill this in: 

## Forcasting model

Use a model to forecast the electricity demand (with **prediction** intervals) that you would expect for the next day if the maximum temperature was $15^\circ$. Compare this with the forecast if the with maximum temperature was $35^\circ$. Do you believe these forecasts? Why or why not? 
 
```{r produce a forecast with the model}
#Fill this in

```

> 'Fill this in:' 


