# (12 points) Vector autoregression

```{r load VAR packages, message=FALSE,message=FALSE}
library(tidyverse)
```

Annual values for real mortgage credit (RMC), real consumer credit (RCC) and real disposable personal income (RDPI) for the period 1946-2006 are recorded in `./data/mortgage_credit.csv`. 

All of the observations are measured in billions of dollars, after adjustment by the Consumer Price Index (CPI). 

Our goal is to develop a VAR model for these data for the period 1946-2003, and then forecast the last three years, 2004-2006. 


```{r read credit data, message=FALSE}
credit <- read_csv('./data/mortgage_credit.csv')
#glimpse(credit)
```

## Time series plot

Plot the time-series of real mortgage credit (RMC), real consumer credit (RCC) and real disposable personal income (RDPI)? Do they look stationary?

```{r Time series plot}
#Fill this in

```

> 'Fill this in


## Check for the unit root

Plot ACF/PACF and Perform the unit root test on these variables and report the results. Do you reject the null of unit root for them? Is the first differencing necessary? 

```{r ACF/PACF and  unit root test}
#Fill this in

```

> 'Fill this in


## Determine VAR model

Based on the unit root results transform the variables and determine the lag length of the VAR using the information criteria.

```{r Determine VAR model}
#Fill this in

```

> 'Fill this in


## Estimation 

Estimate the selected VAR in previous part and comment on the results.

```{r Estimate VAR}
#Fill this in

```

> 'Fill this in


## Model diagnostic

Do diagnostic checking of the VAR model.

```{r Model diagnostic}
#Fill this in

```

> 'Fill this in


## Forecasting

forecast the last three years, 2004-2006.


```{r Forecasting}
#Fill this in

```

> 'Fill this in
